---
title: "Predicting White-throated Sparrow Sex by Wing Chord Length"
author: "Daniel Redwine"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 1  # up to one depth of headings, no subheader (specified by # in html)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme
---
```{r setup, echo=FALSE, warning=FALSE, include = FALSE}
knitr::opts_chunk$set(fig.asp = .66, fig.width = 5)
#Its common to have a setup area to help keep track of required packages and formatting options

library(tidyverse)  #importing, tidying, plotting data
library(knitr)      #making tables
library(leaflet)
library(dplyr)
library(hexbin)
library(ggplot2)
library(readr)
library(RColorBrewer)
library(ggfortify)
library(tidyr)
library(gridExtra)
library(gtsummary)
library(gt)
library(flextable)
library(arm)
library(performance)
library(AICcmodavg)
library(grid)

#knitr::opts_chunk$set(fig.width = 5, fig.asp = 1/3) #force figures to be a certain size and aspect ratio
#opts_chunk$set(echo=FALSE)   #can force hide all code in the output; disables 'code folding'
```
# Sex and Wing Chord Length

Previously the significant difference in wing chord length between males and females has been demonstrated. A GLM with Binomial distribution will be used to demonstrate if wing chord length can be used to predict sex. First we will visualize the raw data. The factors male and female have been transformed to values of 1 and 0 respectively. 

<center>
<iframe src="https://macaulaylibrary.org/asset/155039781/embed" height="515" width="640" frameborder="0" allowfullscreen></iframe>
</center>

```{r Load and clean dataset, echo=FALSE, message=FALSE, warning=FALSE}


wtsp_data <- read.csv("../data/wtsp_data.csv") #load in dataset, .. moves up one level
wing_sex_clean <- wtsp_data %>%
  dplyr::filter(New.Recap == "N") %>% 
  # Using dplyr::filter because multiple packages have a filter command
  # Remove recaps so no repeated measures which would violate assumptions for stats
  dplyr::select(PCRsex, Wing, PCRMorph) %>%
  dplyr::filter(PCRMorph == "WS" | PCRMorph == "TS" ) %>%
  dplyr::filter((PCRsex == "M" & Wing >= 68) | (PCRsex == "F" & Wing <= 74))
  # Selecting only the variables or column headers that I want to use and removing outliers


  wing_sex_clean$PCRsex <- as.factor(wing_sex_clean$PCRsex) #Making sure factors are factors
  wing_sex_clean$Wing <- as.numeric(wing_sex_clean$Wing)
  wing_sex_clean$PCRsex <- as.numeric(factor(wing_sex_clean$PCRse) ) -1
  wing_sex_clean$PCRMorph <- as.factor(wing_sex_clean$PCRMorph)
  

## Needed to clean up the data. Removed all instances of blank or undesired characters in the dataset, changed variables to factors or numeric

```

# Modelling Sex by Wing Chord Length

First we will visualize the data using geom_smooth with glm method.

```{r Male probability vs wing chord graph, warning=FALSE, echo=FALSE, message=FALSE, fig.show = "hold", out.width = "85%", fig.align = "center", fig.cap= "Figure 1. Probability of a White-throated Sparrow being male based on wing chord length. Points at 1 and 0 represent male and female wing chord measurements respectively while the line represents probability of a given measurement belonging to a male bird"}

wing_sex_prob <- ggplot(wing_sex_clean, aes(Wing, PCRsex)) +
  geom_count(show.legend = FALSE) + # Geom count changes size of points to count
  geom_smooth(method="glm", method.args=list(family="binomial"(link="logit")), color = "slateblue", fill = "lightskyblue2") + # Geom smooth can use the model to create curve and CI
  ylab ("Male Probability") +
  xlab ("Wing Chord Length (mm)") +
  theme_bw()
# Create a graph with male/female (1/0) on y and wing chord on x 

wing_sex_prob #Call object 

```
Next we will create the model, a GLM with binomial distribution with a predictor of wing chord length for the response of sex. 

```{r wing_chord_glm, echo=FALSE, message=FALSE, warning=FALSE}

model_sex_wing <- glm(PCRsex ~ Wing, data=wing_sex_clean, binomial(link = 'logit'))
# Fit model, sex as function of wing length, binomial with logit link

model_sex_wing


```
We will then check for overdispertion using library arm. 
```{r sex by wing model assumptions, echo=FALSE, warning=FALSE, message=FALSE, fig.show = "hold", out.width = "85%", fig.align = "center", fig.cap= "Figure 2. Checking overdispersion of data. Note that most points fall outside the 95% CI"}

library(arm)
x <- predict(model_sex_wing) #set x and y
y <- resid(model_sex_wing)
wing_sex_assumption_plot <- binnedplot(x, y) # Create the plot to check confidence intervals

wing_sex_assumption_plot
```
# Interpreting the Model

To interpret results we will use coef and confint.
```{r wtsp sex wing model coefficients, message=FALSE, warning=FALSE, echo=FALSE}

coef(model_sex_wing)

```
The coefficient for wing is 1.927. Using the divide by four method each 1mm increase in wing chord length leads to 48.2% increase in probability that a bird is a male.

```{r wtsp sex wing model CI, message=FALSE, warning=FALSE, echo=FALSE}

confint(model_sex_wing)
```

Here we can see the confidence intervals. 

```{r sex wing model summary, message=FALSE, warning=FALSE, echo=FALSE}
summary(model_sex_wing)
```

Summary of the model predicting sex by wing chord length.